### 2020-12-13 latex/acrobot.tex
First full edit of latest acrobot proof

These dynamics do not always gain energy; rather, energy gain is only
guaranteed under certain conditions on the size of \(I\).
First, let 
\[
    E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
    ,
\]
be the energy function obtained by setting \(I = 0\).
As we will see later, this is the mechanical energy of a simple pendulum with
two masses.
Next, define the set
\begin{equation}\label{eqn:oscillation-domain}
    \mathcal{O}_1 := \left\{(q_u,p_u) \in \SxR 
    \mid E(q_u,p_u) < E(\pi,0) \right\}
    ,
\end{equation}
which is seen in Figure \ref{fig:acrobot-oscillation-domain}.
Finally, let \(\bar{\rho} > \sqrt{60m^2gl^3}\) and define the set
\begin{equation}\label{eqn:o-rhobar}
    \mathcal{O}_2(\bar{\rho}) := \left\{(q_u,p_u) \in \SxR
        \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    ,
\end{equation}
as in Figure \ref{fig:acrobot-o2}

@@

\begin{thm}\label{thm:acrobot-energy-stabilization}
    Consider the simple acrobot ~\eqref{eqn:acrobot-hamiltonian} constrained by
    the VNHC ~\eqref{eqn:acrobot-constraint}, whose
    constraint manifold is \(\Gamma \simeq \SxR\).
\begin{enumerate}
    \item There exists \(I > 0\) small enough that 
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
    \(\mathcal{O}_1\).
    What's more, almost every orbit will escape the closure of 
    \(\mathcal{O}_1\) and begin rotating.
    If instead \(I < 0\), the VNHC dissipates energy on this set.
\item Define \(b : \SxR_{> 0} \rightarrow \R\) by
    \[
        b(\beta,\rho_0) := 
        \frac{5m^2 g l^3 \bar{q}_a \left(
            m^2gl^3\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
            - c_\beta\rho_0^2
        \right)}{
        |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
        }
        ,
    \]
    and set \(S(\rho_0) := \int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma\).
    Fix \(\bar{\rho} > \sqrt{60m^2gl^3}\).
    Suppose there exists \(\epsilon > 0\) so that \(S(\rho_0) > \epsilon\) for
    all \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
    Then there exists \(I > 0\) small enough that
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
    \(\mathcal{O}_2(\bar{\rho})\).
    If instead \(I < 0\), the VNHC dissipates energy on this set.
\end{enumerate}
\end{thm}

Notice that \(\mathcal{O}_1 \subset \mathcal{O}_2(\bar{\rho})\), yet
Theorem \ref{thm:acrobot-energy-stabilization} considers these sets separately.
This separation is advantageous: 
regardless of one's choice of acrobot, 
a small enough fixed value \(I > 0\) will enable the acrobot to 
escape any compact subset of \(\mathcal{O}_1\) in finite time. 
What's more, the acrobot's phase will flow towards (and through) the homoclinic orbit
corresponding to \(E(\pi,0)\), as in Figure
\ref{fig:acrobot-oscillation-domain}.

Hence, the first result of Theorem \ref{thm:acrobot-energy-stabilization} states
that all acrobots constrained by ~\eqref{eqn:acrobot-constraint} 
will gain enough energy to begin rotating around the bar.
In the worst case, the acrobot will at least perform a swing-up routine to reach
the unstable equilibrium at \((q_u,p_u) = (\pi,0)\).

The second result of Theorem
\ref{thm:acrobot-energy-stabilization} pertains to energy injection once the
acrobot has started rotating - that is, when its phase lies in 
\(\mathcal{O}_2(\bar{\rho}) \backslash \mathcal{O}_1\).
If we want the acrobot to achieve giants with energy
\(E(0,\bar{\rho})\), the acrobot needs to satisfy the assumption on the
integral of \(b(\beta,\rho_0)\).
In this case, orbits will escape compact subsets of
\(\mathcal{O}_2(\bar{\rho})\) in finite time until the acrobot eventually
reaches a momentum of at least \(\bar{\rho}\).
One example of this behaviour is displayed in Figure \ref{fig:acrobot-o2}.
Note that the value of \(I\) will depend on \(\bar{\rho}\).

Unfortunately, the proof used in Chapter \ref{ch:vlp} does not
readily transfer to the acrobot.
Proving Theorem \ref{thm:acrobot-energy-stabilization} requires an intelligent
change of coordinates and the use perturbation
theory \cite{khalil_nonlinear}.
We provide the full proof of this theorem in Chapter \ref{sec:acrobot-proof}.

\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:acrobot-proof}
In an effort to make the proof of Theorem \ref{thm:acrobot-energy-stabilization}
as clear as possible, we will break it down into the following segments:
\begin{enumerate}
    \item Background on perturbation theory.
    \item Perturbation analysis for oscillations.
    \item Perturbation analysis for rotations.
\end{enumerate}

@@

\subsection{Perturbation Analysis for Oscillations}
As we have seen, setting \(I = 0\) turns the acrobot into a nominal pendulum. 
This pendulum oscillates whenever the nominal energy
~\eqref{eqn:acrobot-nominal-E} is less than \(E(\pi,0)\). 
Hence, the domain of oscillations for the nominal pendulum is given by
\[
    \mathcal{O}_1 := \left\{ (q_u,p_u) \in \SxR \mid E(q_u,p_u) < E(\pi,0)\right\}
    .
\]
Recall that orbits of a pendulum are level sets of \(E\). 
From Figure \ref{fig:pendulum-level-sets}, we know that oscillation level sets
look like ellipses contained in \(\mathcal{O}_1\).
Since perturbation theory requires us to analyze the nominal pendulum, we will
make our analysis convenient by changing coordinates into a
pseudo-radius \(\mu > 0\) which remains constant on level sets of \(E\), 
along with a pseudo-angle \(\alpha \in \Sone\) satisfying \(\dot{\alpha} > 0\) on
\(\mathcal{O}\).

@@

We have thus found a transformation \(T\inv(\alpha,\mu) = (q_u,p_u)\).
We need the inverse of this map to get our diffeomorphism \(T(q_u,p_u)\).
Notice from ~\eqref{eqn:oscillation-pu2} that
\[
    \cos(\mu) = -\frac{p_u^2}{30m^2gl^3} + \cos(q_u) =: C_\mu(q_u,p_u)
    .
\]
Since \(\mu \in ]0,\pi[\), we can uniquely express \(\mu\) by
\begin{equation*}
    \mu = \arccos\left(C_\mu(q_u,p_u)\right)
    .
\end{equation*}
Next we need to find \(\alpha\). 
Recall that 
\begin{equation}\label{eqn:acrobot-cosalpha-qu}
    \cos(\alpha) = \frac{q_u}{\mu}
    ,
\end{equation}
which means 
\begin{equation}\label{eqn:acrobot-sinalpha-qu}
    \sin(\alpha) = \pm \sqrt{1 - \frac{q_u^2}{\mu^2}}
    .
\end{equation}
Using ~\eqref{eqn:oscillation-pu}, we determine that
\(\sign{\sin(\alpha)} = -\sign{p_u}\).
Putting together ~\eqref{eqn:acrobot-cosalpha-qu} and
~\eqref{eqn:acrobot-sinalpha-qu}, we deduce that
\begin{equation*}
    \alpha = \left.
        \arctan_2\left( -\sign{p_u}\sqrt{1 - \frac{q_u^2}{\mu^2}}, \frac{q_u}{\mu}\right)
        \right|_{\mu = \arccos(C_\mu(q_u,p_u))}
    .
\end{equation*}
Thus, our transformation into \((\alpha,\mu)\)-coordinates is
\begin{align}
        \label{eqn:alpha-qupu}
        \alpha &= \left.
        \arctan_2\left( -\sign{p_u}\sqrt{1 - \frac{q_u^2}{\mu^2}}, \frac{q_u}{\mu}\right)
        \right|_{\mu = \arccos(C_\mu(q_u,p_u))}
        ,\\
        \label{eqn:mu-qupu}
        \mu &= \arccos\left(C_\mu(q_u,p_u)\right).
        .
\end{align}

@@

Since \(\dot{\alpha} > 0\), there is a well-defined Poincar\'{e} map 
describing how the pseudo-radius \(\mu\) changes each time the orbit of \((q_u(t),p_u(t))\)
intersects the \(q_u\)-axis, \ie, when \(\alpha = 0\) or \(\alpha = \pi\).
Using our time-scaled system ~\eqref{eqn:muhat-dot}, we can define this
Poincar\'{e} map as the change in \(\hat{\mu}\) when \(\alpha\) increases by
\(\pi\):
\begin{align*}
    P_\mathcal{O} : \,]0,\pi[ &\rightarrow \R_+,
    \\
    \mu_0 &\mapsto \hat{\mu}(\pi,\mu_0,I)
    .
\end{align*}
The Poincar\'{e} map expands into
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + I K \int\limits_0^\pi a(\sigma,\mu_0)d\sigma
    + R(\pi,\mu_0,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi a(\sigma,\mu_0)d\sigma
    .
\]
Note that \(K\) is a positive constant which contains the acrobot's physical
parameters \(m\), \(g\), \(l\), and \(\bar{q}_a\), 
while \(a(\alpha,\mu)\) is adimensional. 
This means \(Q(\mu_0)\) is identical for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive for any acrobot.

@@

A sufficient conditions for this characterization is to prove that 
\(P_\mathcal{O}(\mu_0) \geq \mu_0 + \lambda\) for some \(\lambda > 0\).
We will separate our analysis of the Poincar\'{e} map into two sections.
First, we will prove that orbits near the origin flow away from it by showing the
origin is a repeller of ~\eqref{eqn:muhat-discrete}.
Then, we will use the sufficient condition to guarantee all orbits reach 
\(\mu = \pi\).

Linearizing ~\eqref{eqn:muhat-discrete} at \(\mu_0 = 0\) yields
\[
    P^\prime_\mathcal{O}(0) = 1 + I K Q^\prime(0) +
    R^\prime(\pi,0,I)
    ,
\]
where prime denotes differentiation with respect to \(\mu\).
If there is some value of \(I \in \,]0,I_1]\) for which this term is greater than
\(1\), then \(0\) is a repeller of the discrete time system.
To show such a value of \(I\) exists, we first compute 
\[
    Q^\prime(0) = \int \limits_0^\pi \pdiff{a}{\mu}(\sigma,0) d\sigma
    .
\]
Numerical computations reveal that 
\[
    \lim \limits_{\mu_0 \to 0^+}
    \pdiff{a}{\mu}(\alpha,\mu_0) 
    = -\frac{\sqrt{2}}{2} \left(11\sin(\alpha)^2 - 6\right)
    ,
\]
which means 
\[
    Q^\prime(0)  = \frac{\pi}{2\sqrt{2}} > 0
    .
\]
Since the remainder term \(R(\pi,0,I)\) is \(O(I^2)\), 
its partial derivative \(R^\prime(\pi,0,I)\) is also \(O(I^2)\).
Hence, it can be written in the form
\(I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and
zero at \(I = 0\).
Thus, there exists \(I_2 \in \,]0,I_1]\) where 
\(I_2 K Q^\prime(0) + (I_2)^2 \tilde{R}(I) > 0\).
Hence,
\[
    P^\prime_\mathcal{O}(0) \geq 1 + IKQ^\prime(0) + I^2 \tilde{R}(I) 
    > 1
    ,
\] 
for all \(I \in \,]0, I_2]\).

@@

To complete this proof, recall that \(R(\pi,\mu_0,I)\) is smooth in all its
parameters as well as being \(O(I^2)\).
It is therefore bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
by some value \(\underbar{R}(I)\). 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some
\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in [-I_3, I_3]\),
\[
    R(\pi,\mu_0,I) \geq \underbar{R}(I) > -I^2 r
    .
\]
Note that we can assume \(I_3 \leq I_2\) without loss of generality.
Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
for \(\mu_0 \in [\epsilon,\pi[\),
\[
    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
    .
\]
Picking a small \(\lambda > 0\) and choosing 
\(I^\star \in\, ]0, I_3]\) so that
\[
    I^\star K Q(\epsilon) - (I^\star)^2 r \geq \lambda > 0
    ,
\]
means that \( P_\mathcal{O}(\mu_0) \geq \mu_0 + \lambda\)
for all \(\mu_0 \in [\epsilon,\pi[\).

Looking at ~\eqref{eqn:muhat-discrete} we find that
\(\mu_{n+1} \geq \mu_n + \lambda\). 
This implies that all solutions of the discrete time system will flow towards
\(\mu = \pi\).

@@

By the same arguments presented in this section, the Poincar\'{e} section
satisfies \( P_\mathcal{O}(\mu_0) \leq \mu_0 - \lambda\) when \(I < 0\). 
In this case, the constraint dissipates energy on \(\mathcal{O}_1\).

We have proven the first part of Theorem \ref{thm:acrobot-energy-stabilization},
which claims there exists \(I > 0\) small enough that 
~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
\(\mathcal{O}_1\).

Here is a summary of the proof:
\begin{enumerate}
    \item We defined pseudo-polar coordinates \((\alpha,\mu)\) adapted to level
        sets of the nominal pendulum's mechanical energy.
    \item We showed there exists a value \(I_1\) small enough where 
        \(\dot{\alpha} > 0\) everywhere on \(\mathcal{O}_1\).
    \item We used time scaling to find \(\hat{\mu}\) and
        expanded it into ~\eqref{eqn:acrobot-muhat-approx} using
        perturbation theory.
    \item Taking this expanded solution, we defined and expanded the
        Poincar\'{e} map \(P_\mathcal{O}\) to analyze the discrete-time system
        ~\eqref{eqn:muhat-discrete}.
    \item We found a value \(I_2 \leq I_1\) making 0 a repeller of the
        ~\eqref{eqn:muhat-discrete}, and a value \(I_3 \leq I_2\) which drives
        all its orbits towards \(\mu = \pi\).
\end{enumerate}
Hence, choosing \(I \in \,]0,I_3]\) guarantees energy injection on
\(\mathcal{O}_1\) because all orbits approach the pseudo-radius \(\mu = \pi\).
Likewise, choosing \(I \in [-I_3,0[\) guarantees energy dissipation on
\(\mathcal{O}_1\).

\subsection{Energy Gain on \(\bar{\mathcal{O}}_1\)}
Before moving on to the rotation analysis, we must first confirm that any
acrobot constrained by ~\eqref{eqn:acrobot-constraint} will eventually start
rotating.
The definition of energy gain states that all orbits initialized in
\(\mathcal{O}_1\) will escape compact sets of \(\mathcal{O}_1\) in finite time. 
This is not enough to prove that all orbits will escape the boundary of
\(\mathcal{O}_1\).

@@

Using \(x(t) := (q_u(t),p_u(t))\) as shorthand, let
\(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
Suppose the orbit \(x(\R)\) does not cross through \(E_\pi\)
into the rotation zone.
Since the acrobot gains energy on \(\mathcal{O}_1\), the positive limit set of
\(x(t)\) must be \(E_\pi\). 
By the Birkhoff Theorem, \(x(t)\) converges to an invariant subset of \(E_\pi\)
\cite{birkhoff}.
From our discussion on pseudo-polar coordinates (in particular the extension
~\eqref{eqn:alpha-dot-boundary} of \(\dot{\alpha}\) to \(\bar{\mathcal{O}}_1\)),
we know that \(\dot{\alpha} \geq 0\) on \(E_\pi\), with equality if and only if 
\(\alpha \in \{0,\pi\}\).
In other words, the only invariant subset of \(E_\pi\) is the upright equilibrium.
Hence, \(x(t)\) must converge to \((\pi,0)\), which means \(x(0) \in \Pi^+\).
Since we know that \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every
other orbit initialized in \(\mathcal{O}_1\) must escape through the boundary
\(E_\pi\) into the rotation domain.

\subsection{Perturbation Analysis for Rotations}
Figure \ref{fig:pendulum-level-sets} reminds us that rotations of the 
nominal pendulum obtained by setting \(I = 0\) look like open curves on some
rotation region \(\mathcal{R} \subset \SxR\).
One can compute this rotation domain to be the set
\[
    \mathcal{R} := \left\{ (q_u,p_u) \in \SxR \mid E(q_u,p_u) > E(\pi,0)\right\}
    .
\]
Performing a similar process to what we did for oscillations,
we wish to find a new set of coordinates \((\beta,\rho)\)
where the pseudo-radius \(\rho \in \mathbb{R}\) remains constant on level sets
of \(E\) and the pseudo-angle \(\beta \in \Sone\) is always increasing on
\(\mathcal{R}\).

>>>>

These dynamics do not always gain energy; rather, energy gain is only
guaranteed under certain conditions on the size of \(I\).
First, let 
\[
    E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
    ,
\]
be the energy function obtained by setting \(I = 0\).
As we will see later, this is the mechanical energy of a simple pendulum with
two masses.
Next, define the set
\begin{equation}\label{eqn:oscillation-domain}
    \mathcal{O}_1 := \left\{(q_u,p_u) \in \SxR 
    \mid E(q_u,p_u) < E(\pi,0) \right\}
    ,
\end{equation}
as in Figure \ref{fig:acrobot-oscillation-domain}.
Finally, let \(\bar{\rho} > \sqrt{60m^2gl^3}\) and define the set
\begin{equation}\label{eqn:o-rhobar}
    \mathcal{O}_2(\bar{\rho}) := \left\{(q_u,p_u) \in \SxR
        \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    ,
\end{equation}
as in Figure \ref{fig:acrobot-o2}

@@

\begin{thm}\label{thm:acrobot-energy-stabilization}
    Consider the simple acrobot ~\eqref{eqn:acrobot-hamiltonian} constrained by
    the VNHC ~\eqref{eqn:acrobot-constraint}, whose
    constraint manifold is \(\Gamma \simeq \SxR\).
\begin{enumerate}
    \item There exists \(I > 0\) small enough that 
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
    \(\mathcal{O}_1\).
    What's more, almost every orbit will escape the closure of 
    \(\mathcal{O}_1\).
    If instead \(I < 0\), the VNHC dissipates energy.
\item Define \(b : \SxR_{> 0} \rightarrow \R\) by
    \[
        b(\beta,\rho_0) := 
        \frac{5m^2 g l^3 \bar{q}_a \left(
            m^2gl^3\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
            - c_\beta\rho_0^2
        \right)}{
        |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
        }
        ,
    \]
    and define \(S(\rho_0) := \int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma\).
    Fix \(\bar{\rho} > \sqrt{60m^2gl^3}\).
    Suppose there exists \(\epsilon > 0\) so that \(S(\rho_0) \geq \epsilon\) for
    all \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
    Then there exists \(I > 0\) small enough that
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
    \(\mathcal{O}_2(\bar{\rho})\).
    If instead \(I < 0\), the VNHC dissipates energy.
\end{enumerate}
\end{thm}

Notice that \(\mathcal{O}_1 \subset \mathcal{O}_2(\bar{\rho})\), yet
Theorem \ref{thm:acrobot-energy-stabilization} considers these sets separately.
This separation is advantageous because the first result holds for any
\(m\), \(g\), \(l\), and \(\bar{q}_a\). 
That is, regardless of one's choice of acrobot, 
a small enough fixed value \(I > 0\) will enable the acrobot to 
escape any compact subset of \(\mathcal{O}_1\) in finite time. 
What's more, the acrobot's phase will flow towards (and through) the homoclinic orbit
corresponding to \(E(\pi,0)\), as in Figure
\ref{fig:acrobot-oscillation-domain}.

In other words, the first result of Theorem
\ref{thm:acrobot-energy-stabilization} states that all acrobots constrained by
~\eqref{eqn:acrobot-constraint} will gain enough energy to begin rotating around
the bar.
In the worst case, the acrobot will at least perform a swing-up routine to reach
the unstable equilibrium at \((q_u,p_u) = (\pi,0)\).

The second result of Theorem
\ref{thm:acrobot-energy-stabilization} pertains to energy injection once the
acrobot has started rotating.
For the acrobot to achieve giants with energy
\(E(0,\bar{\rho})\), it must satisfy the assumption on the integral of
\(b(\beta,\rho_0)\).
The value of this integral depends on the acrobot's physical parameters.
If the assumption holds, orbits will escape compact subsets of
\(\mathcal{O}_2(\bar{\rho})\) in finite time until the acrobot eventually
reaches a momentum of at least \(\bar{\rho}\).
One example of this behaviour is displayed in Figure \ref{fig:acrobot-o2}.
Note that the size of \(I\) will depend on \(\bar{\rho}\).

The proof used in Chapter \ref{ch:vlp} does not
readily transfer to the acrobot.
Proving Theorem \ref{thm:acrobot-energy-stabilization} requires an intelligent
change of coordinates and the use perturbation
theory \cite{khalil_nonlinear}.
We provide the full proof of this theorem in Chapter \ref{sec:acrobot-proof}.

\section{Proof of Theorem \ref{thm:acrobot-energy-stabilization}}\label{sec:acrobot-proof}
To make the proof of Theorem \ref{thm:acrobot-energy-stabilization}
as clear as possible, we break it down into the following segments:
\begin{enumerate}
    \item Background on perturbation theory.
    \item Perturbation analysis for oscillations.
    \item Perturbation analysis for rotations.
\end{enumerate}

@@

\subsection{Perturbation Analysis for Oscillations}\label{sec:acrobot-proof-o1}
As we have seen, setting \(I = 0\) turns the acrobot into a nominal pendulum. 
This pendulum oscillates whenever the nominal energy
~\eqref{eqn:acrobot-nominal-E} is less than \(E(\pi,0)\). 
Hence, the domain of oscillations for the nominal pendulum is given by
\[
    \mathcal{O}_1 := \left\{ (q_u,p_u) \in \SxR \mid E(q_u,p_u) < E(\pi,0)\right\}
    .
\]
Recall that orbits of a pendulum are level sets of \(E\). 
From Figure \ref{fig:pendulum-level-sets}, we know that oscillation level sets
look like ellipses contained in \(\mathcal{O}_1\).
Since perturbation theory requires us to analyze the nominal pendulum, we will
make our analysis convenient by changing coordinates into a
pseudo-radius \(\mu > 0\) which remains constant on level sets of \(E\), 
along with a pseudo-angle \(\alpha \in \Sone\) satisfying \(\dot{\alpha} > 0\) on
\(\mathcal{O}_1\).

@@

We have thus found a transformation \(T\inv(\alpha,\mu) = (q_u,p_u)\).
We need the inverse of this map to get our diffeomorphism \(T(q_u,p_u)\).
Notice from ~\eqref{eqn:oscillation-pu2} that
\[
    \cos(\mu) = -\frac{p_u^2}{30m^2gl^3} + \cos(q_u) =: C_\mu(q_u,p_u)
    .
\]
Since \(\mu \in ]0,\pi[\), we can uniquely express \(\mu\) by
\begin{equation*}
    \mu = \arccos\left(C_\mu(q_u,p_u)\right)
    .
\end{equation*}
Next we need to find \(\alpha\). 
Recall that 
\begin{equation}\label{eqn:acrobot-cosalpha-qu}
    \cos(\alpha) = \frac{q_u}{\mu}
    ,
\end{equation}
which means 
\begin{equation}\label{eqn:acrobot-sinalpha-qu}
    \sin(\alpha) = \pm \sqrt{1 - \frac{q_u^2}{\mu^2}}
    .
\end{equation}
Using ~\eqref{eqn:oscillation-pu}, we determine that
\(\sign{\sin(\alpha)} = -\sign{p_u}\).
Putting together ~\eqref{eqn:acrobot-cosalpha-qu} and
~\eqref{eqn:acrobot-sinalpha-qu}, we deduce that
\begin{equation*}
    \alpha = \left.
        \arctan_2\left( -\sign{p_u}\sqrt{1 - \frac{q_u^2}{\mu^2}}, \frac{q_u}{\mu}\right)
        \right|_{\mu = \arccos(C_\mu(q_u,p_u))}
    .
\end{equation*}
Thus, our transformation into \((\alpha,\mu)\)-coordinates is
\begin{align}
        \label{eqn:alpha-qupu}
        \alpha &= \left.
        \arctan_2\left( -\sign{p_u}\sqrt{1 - \frac{q_u^2}{\mu^2}}, \frac{q_u}{\mu}\right)
        \right|_{\mu = \arccos(C_\mu(q_u,p_u))}
        ,\\
        \label{eqn:mu-qupu}
        \mu &= \arccos\left(C_\mu(q_u,p_u)\right)
        .
\end{align}

@@

Since \(\dot{\alpha} > 0\), there is a well-defined Poincar\'{e} map 
describing how the pseudo-radius \(\mu\) changes each time the orbit of \((q_u(t),p_u(t))\)
intersects the \(q_u\)-axis, \ie, when \(\alpha = 0\) or \(\alpha = \pi\).
Using our time-scaled system ~\eqref{eqn:muhat-dot}, we can define this
Poincar\'{e} map as the change in \(\hat{\mu}\) when \(\alpha\) increases by
\(\pi\):
\begin{align*}
    P_\mathcal{O} : \,]0,\pi[ &\rightarrow \R_+,
    \\
    \mu_0 &\mapsto \hat{\mu}(\pi,\mu_0,I)
    .
\end{align*}
The Poincar\'{e} map expands into
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + I K \int\limits_0^\pi a(\sigma,\mu_0)d\sigma
    + R(\pi,\mu_0,I)
    .
\]
Let us define
\begin{equation}\label{eqn:acrobot-Qmu}
    Q(\mu_0) := \int \limits_0^\pi a(\sigma,\mu_0)d\sigma
    .
\end{equation}
Note that \(K\) is a positive constant which contains the acrobot's physical
parameters \(m\), \(g\), \(l\), and \(\bar{q}_a\), 
while \(a(\alpha,\mu)\) is adimensional. 
This means \(Q(\mu_0)\) is identical for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive for any acrobot.

@@

A sufficient conditions for this characterization is to prove that 
\(P_\mathcal{O}(\mu_0) \geq \mu_0 + \gamma\) for some \(\gamma > 0\).
We will perform our analysis in two sections.
First, we will show the origin is a repeller of ~\eqref{eqn:muhat-discrete}.
This implies that orbits near the origin flow away from it.
Then, we will use the sufficient condition above to guarantee all orbits reach
\(\mu = \pi\).

Linearizing ~\eqref{eqn:muhat-discrete} at \(\mu_0 = 0\) yields
\[
    P^\prime_\mathcal{O}(0) = 1 + I K Q^\prime(0) +
    R^\prime(\pi,0,I)
    ,
\]
where prime denotes differentiation with respect to \(\mu\).
If there is some value of \(I \in \,]0,I_1]\) for which this term is greater than
\(1\), then \(0\) is a repeller of the discrete time system.
To show such a value of \(I\) exists, we first compute 
\[
    Q^\prime(0) = \int \limits_0^\pi \pdiff{a}{\mu}(\sigma,0) d\sigma
    .
\]
Numerical computations reveal that 
\[
    \lim \limits_{\mu_0 \to 0^+}
    \pdiff{a}{\mu}(\alpha,\mu_0) 
    = -\frac{\sqrt{2}}{2} \left(11\sin(\alpha)^2 - 6\right)
    ,
\]
which means 
\[
    Q^\prime(0)  = \frac{\pi}{2\sqrt{2}} > 0
    .
\]
Since the remainder term \(R(\pi,0,I)\) is \(O(I^2)\), 
its partial derivative \(R^\prime(\pi,0,I)\) is also \(O(I^2)\).
Hence, it can be written in the form
\(I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and
zero at \(I = 0\).
Thus, there exists \(I_2 \in \,]0,I_1]\) such that
\[
    I_2 K Q^\prime(0) + (I_2)^2 \tilde{R}(I) > 0
    .
\]
Hence,
\[
    P^\prime_\mathcal{O}(0) \geq 1 + IKQ^\prime(0) + I^2 \tilde{R}(I) 
    > 1
    ,
\] 
for all \(I \in \,]0, I_2]\).

@@

To complete the proof, recall that \(R(\pi,\mu_0,I)\) is smooth in all its
parameters as well as being \(O(I^2)\).
It is therefore bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
by some value \(\underbar{R}(I)\). 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some
\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in [-I_3, I_3]\),
\[
    R(\pi,\mu_0,I) \geq \underbar{R}(I) > -I^2 r
    .
\]
Note that we can assume \(I_3 \leq I_2\) without loss of generality.
Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
for \(\mu_0 \in [\epsilon,\pi[\),
\[
    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
    .
\]
Picking a small \(\gamma > 0\) and choosing 
\(I^\star \in\, ]0, I_3]\) so that
\[
    I^\star K Q(\epsilon) - (I^\star)^2 r \geq \gamma > 0
    ,
\]
means that \( P_\mathcal{O}(\mu_0) \geq \mu_0 + \gamma\)
for all \(\mu_0 \in [\epsilon,\pi[\).

Looking at ~\eqref{eqn:muhat-discrete} we find that
\(\mu_{n+1} \geq \mu_n + \gamma\). 
This implies that all solutions of the discrete time system will flow towards
\(\mu = \pi\).

@@

By the same arguments presented in this section, the Poincar\'{e} section
satisfies \( P_\mathcal{O}(\mu_0) \leq \mu_0 - \gamma\) when \(I < 0\). 
In this case, the constraint dissipates energy on \(\mathcal{O}_1\).

\subsubsection*{Energy Gain on \(\bar{\mathcal{O}}_1\)}
Before moving on to the rotation analysis, we must first confirm that any
acrobot constrained by ~\eqref{eqn:acrobot-constraint} will eventually start
rotating.
The definition of energy gain states that all orbits initialized in
\(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite time. 
This is not enough to prove that all orbits will escape the boundary of
\(\mathcal{O}_1\).

@@

Using \(x(t) := (q_u(t),p_u(t))\) as shorthand, let
\(x(0) \in \mathcal{O}_1\) be a nonzero initial condition of the acrobot.
Suppose the orbit \(x(\R)\) does not cross through \(E_\pi\)
into the rotation zone.
Since the acrobot gains energy on \(\mathcal{O}_1\), the positive limit set of
\(x(t)\) must be \(E_\pi\). 
By the Birkhoff Theorem \cite{birkhoff}, \(x(t)\) converges to an invariant
subset of \(E_\pi\).
From our discussion on pseudo-polar coordinates (in particular the extension
~\eqref{eqn:alpha-dot-boundary} of \(\dot{\alpha}\) to \(\bar{\mathcal{O}}_1\)),
we know that \(\dot{\alpha} \geq 0\) on \(E_\pi\), with equality if and only if 
\(\alpha \in \{0,\pi\}\).
In other words, the only invariant subset of \(E_\pi\) is the upright equilibrium.
Hence, \(x(t)\) must converge to \((\pi,0)\), which means \(x(0) \in \Pi^+\).
Since we know that \(\Pi^+\) is a set of measure zero in \(\SxR\), almost every
other orbit initialized in \(\mathcal{O}_1\) must escape through the boundary
\(E_\pi\) into the rotation domain.

We have proven the first part of Theorem \ref{thm:acrobot-energy-stabilization},
which claims there exists \(I > 0\) small enough that 
~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
\(\mathcal{O}_1\).
Enough energy is injected that orbits will exit the closure of \(\mathcal{O}_1\)
and enter the rotation domain.

Here is a summary of the proof:
\begin{enumerate}
    \item We found pseudo-polar coordinates \((\alpha,\mu)\) adapted to level
        sets of the nominal pendulum's mechanical energy on \(\mathcal{O}_1\).
    \item We showed there exists a value \(I_1\) small enough where 
        \(\dot{\alpha} > 0\) everywhere on \(\mathcal{O}_1\).
    \item We found the time-scaled pseudo-radius \(\hat{\mu}\) and
        expanded it into ~\eqref{eqn:acrobot-muhat-approx} using
        perturbation theory.
    \item Taking this expanded solution, we defined and expanded the
        Poincar\'{e} map \(P_\mathcal{O}\) to analyze the discrete-time system
        ~\eqref{eqn:muhat-discrete}.
    \item We found a value \(I_2 \leq I_1\) making the origin a repeller of 
        ~\eqref{eqn:muhat-discrete}, and a value \(I_3 \leq I_2\) which drives
        \(\mu_n\) towards \(\mu = \pi\).
    \item We showed that the upright equilibrium \((\pi,0)\) is unstable, and
        used this fact to prove that almost every orbit crosses through \(E_\pi\)
        into the rotation zone.
\end{enumerate}
Choosing \(I \in \,]0,I_3]\) guarantees energy injection on
\(\mathcal{O}_1\) because all orbits approach the pseudo-radius \(\mu = \pi\).
Likewise, choosing \(I \in [-I_3,0[\) guarantees energy dissipation on
\(\mathcal{O}_1\).

\subsection{Perturbation Analysis for Rotations}
Figure \ref{fig:pendulum-level-sets} reminds us that rotations of the 
nominal pendulum obtained by setting \(I = 0\) look like open curves on some
rotation domain \(\mathcal{R} \subset \SxR\), where
\[
    \mathcal{R} := \left\{ (q_u,p_u) \in \SxR \mid E(q_u,p_u) > E(\pi,0)\right\}
    .
\]
Performing a similar process to what we did for oscillations,
we wish to find a new set of coordinates \((\beta,\rho)\)
where the pseudo-radius \(\rho \in \mathbb{R}\) remains constant on level sets
of \(E\) and the pseudo-angle \(\beta \in \Sone\) is always increasing on
\(\mathcal{R}\).
